---
title: "04 - Immune Trend Visualization"
author: "Jay Nagabhairu"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(dplyr)
library(reshape2)
library(ggplot2)
```

Tracks average immune-cell fractions (from CIBERSORTx, step 03) across the
progression axis, and separately by recurrence status within stage 1.

## Load deconvolution results

```{r load}
dec_dir <- here("results", "deconvolution")

precancer <- read.csv(file.path(dec_dir, "CIBERSORTx_precancer.csv"),
                      stringsAsFactors = FALSE, check.names = FALSE)
stage1    <- read.csv(file.path(dec_dir, "CIBERSORTx_stage1_by_recurrence.csv"),
                      stringsAsFactors = FALSE, check.names = FALSE)
stage2    <- read.csv(file.path(dec_dir, "CIBERSORTx_stage2_by_recurrence.csv"),
                      stringsAsFactors = FALSE, check.names = FALSE)

clin <- read.csv(here("data", "raw", "BC_ClinData_233rows.csv"),
                 stringsAsFactors = FALSE)
```

## Mean fraction of each of the 22 LM22 subsets for a group

```{r means-fn}
# Returns a named numeric vector: mean fraction per immune subset.
# Assumes the first 22 columns after "Mixture" are the LM22 fractions.
funcMeans <- function(cibersortData) {
  frac <- cibersortData[, 2:23]          # 22 LM22 columns (col 1 = Mixture id)
  colMeans(frac, na.rm = TRUE)
}
```

## Progression trajectory: pre-cancer -> stage 1 -> stage 2

```{r progression}
meansProgression <- cbind(
  `Pre-cancer` = funcMeans(precancer),
  `Stage 1`    = funcMeans(stage1),
  `Stage 2`    = funcMeans(stage2)
)
meansProgression <- as.data.frame(meansProgression)
meansProgression$row.names <- rownames(meansProgression)

source(here("R", "functionLineGraph.R"))

fig_dir <- here("docs", "figures")
progressionPlot <- functionLineGraph(
  meansProgression,
  outputFileName   = "immune_progression_trajectory",
  outputFolderName = fig_dir
)
progressionPlot
```

## Recurrence-stratified: stage 1 relapse vs relapse-free

```{r recurrence}
# recurrence: 1 = non-relapse (relapse-free), 2 = relapse
relapseFreeIDs <- clin$GSMid[clin$recurrence == 1]
relapseIDs     <- clin$GSMid[clin$recurrence == 2]

# Filter by matching the CIBERSORTx "Mixture" id (NOT by row index)
stage1RelapseFree <- stage1[stage1$Mixture %in% relapseFreeIDs, ]
stage1Relapse     <- stage1[stage1$Mixture %in% relapseIDs, ]

meansRecurrence <- cbind(
  `Relapse-free` = funcMeans(stage1RelapseFree),
  `Relapse`      = funcMeans(stage1Relapse)
)
meansRecurrence <- as.data.frame(meansRecurrence)
meansRecurrence$row.names <- rownames(meansRecurrence)

recurrencePlot <- functionLineGraph(
  meansRecurrence,
  outputFileName   = "immune_stage1_by_recurrence",
  outputFolderName = fig_dir
)
recurrencePlot
```

Figures are written to `docs/figures/` as PNG, PDF, and TIFF.
